---
title: "Regression Analysis"
author: "Michael Bristow"
date: "10 December 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Regression Models - Peer Assesement 1

##Issue/Problem
Take the mtcars data set and write up an analysis to answer their question using regression models and exploratory data analyses.

Your report must be:

Written as a PDF printout of a compiled (using knitr) R markdown document.

Brief. Roughly the equivalent of 2 pages or less for the main text. 

Supporting figures in an appendix can be included up to 5 total pages including the 2 for the main report. The appendix can only include figures.
    Include a first paragraph executive summary.
    
##Executuve Summary
This report looks at historic data and trys to find drivers for MPG.
It will look at trwo main items  

* Is an automatic or manual transmission better for MPG  
* Quantify the MPG difference between automatic and manual  
 transmissions
 
#Data Processing/Analysis
First load the data
```{r}
library(ggplot2)
data(mtcars)
xx <- mtcars
#convert "am"to a factor
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <-c("AT", "MT")

summary(mtcars)
head(mtcars)

```
Apply regsubsets to find best variable selection

```{r}
library(leaps)
reg.best<- regsubsets(mpg~.,mtcars,nvmax=5)
summary(reg.best)

reg.summary<- summary(reg.best)
names(reg.summary)

```
Look at the best r squared

```{r}
reg.summary$rsq

```
As can be seen, "weight" is the biggest "driver" of mpg followed by number of cylinders. 

Our model should include  

- weight
- cylinders

By adding more variaqbles, the r-squared only increases by small amounts and
we need to be wary of overfitting

Plot weight against mpg and show regression line

```{r}
ggplot(data = xx, aes(x = wt, y = mpg)) + 
  geom_point(color='red') +
  geom_smooth(method = "lm", color = "blue") +
  labs(x = "Vehicle weight") +
  labs(y = "Miles per gallon")
```

As can be seen there is a direct correlation between the vehicle weight and it mpg


```{r}
wtdata <- lm(mpg~wt+cyl, data = mtcars)
summary(wtdata)

```
We can see that each extra ton in weight changes MPG by `r coef(wtdata)[2]` when we use the full model using weight and cylinders


#Effect of Transmission type on MPG
Question is, which has better MPG - Manual or Automatic
We do this by running a regression using ONLY transmission type

```{r}
fit <- lm(mpg~am, data = mtcars)
summary(fit)

```
It shows that on average

- a car has 17.147 mpg with automatic transmission, and if it is manual transmission, 7.245 mpg is increased. 
- has the Residual standard error as 4.902 on 30 degrees of freedom. 
- the Adjusted R-squared value is 0.3385, which means that the model can explain about 34% of the variance of the MPG variable. 
- The low Adjusted R-squared value also indicates that we need to add other variables to the model.
 
```{r}

```
#Conclusion
Here are the conclusions fromn the analysis

- The primary driver of MPG is the cars weight
- The "optimal" model is to inlcude weight and cylinders
- By adding more variables, the r-quared value only increases marginally and we are in danger of overfitting the model
- Transmission typoe also has an efefct on MPG
- Having a manual transmission increases MPG by 7.245